---
title: "class 17 Vaccination Mini Project"
format: pdf
---

# Getting Started

```{r, results='hide', echo = T}
vax <- read.csv("statewide-covid-19-vaccines-administered-by-zip-code.csv")
head(vax)
tail(vax)
```

> Q1. What column details the total number of people fully vaccinated?

persons_fully_vaccinated

> Q2. What column details the Zip code tabulation area?

zip_code_tabulation_area

> Q3. What is the earliest date in this dataset?

2021-01-05

> Q4. What is the latest date in this dataset?

2022-11-22


```{r}
skimr::skim(vax)
```

```{r}
# To find all the NA values in the persons_fully_vaccinated column
sum( is.na(vax$persons_fully_vaccinated) )
```

> Q5. How many numeric columns are in this dataset?

13

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

14921

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

9.2%

Working with Dates

```{r, results='hide'}
library(lubridate)
```


```{r}
today()

# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

Using this format, we can determine the span of the datasets.

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```


> Q9. How many days have passed since the last update of the dataset?

6 Days

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

99 unique dates


# Working with Zip-Codes
```{r, results='hide'}
library(zipcodeR)
```


```{r}
geocode_zip('92037')
```

```{r}
# Calculate distance of zipcode centers in miles

zip_distance('92037','92109')
```

You can pull useful data from zipcodes with this function.

```{r}
reverse_zipcode(c('92037', "92109") )
```

# Focus on the San Diego Area

```{r}
# Subset to San Diego county only areas
sd <- vax[ vax$county == "San Diego" , ]
```

or, use the dplyr package

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)

filter (vax, county == "San Diego" & 
          age12_plus_population > 70000 &
          as_of_date == "2022-11-15")
```

> Q11. How many distinct zip codes are listed for San Diego County?

107

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

92154

```{r}
sd.11 <- filter(vax, county == "San Diego" &
         as_of_date == "2022-11-15")

fullyVaccPercent <- sd.11$percent_of_population_fully_vaccinated

mean(fullyVaccPercent[!is.na(fullyVaccPercent)])
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”?

73.69%

> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?

```{r}
library(ggplot2)

hist(fullyVaccPercent, 
     main = "Histogram of Vaccination Rates Across San Diego County", 
     xlab = "Percent of Population Fully Vaccinated on 2022-11-15")
```

Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")

ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "date", y="Percent Vaccinated", 
       title = "Vaccination rate for La Jolla CA 92109")
```

Comparing to Similar Sized Area

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-11-15")

head(vax.36)
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  geom_hline(yintercept = 0.7172851, linetype = "dashed", col = "red") + 
  labs(x = "date", y="Percent Vaccinated", 
       title = "Vaccination rate for La Jolla CA 92109")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
hist(vax.36$percent_of_population_fully_vaccinated)
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
filter(vax, zip_code_tabulation_area == "92109", as_of_date == "2022-11-15")$percent_of_population_fully_vaccinated
filter(vax, zip_code_tabulation_area == "92040", as_of_date == "2022-11-15")$percent_of_population_fully_vaccinated
```

Both of these area codes are below the average.

```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x = "Date", y = "Percent Vaccinated",
       title = "Vaccination Rate Across California",
       subtitle = "Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.7172851, linetype="dashed", col = "red")
```

> Q21. How do you feel about traveling for Thanksgiving Break and meeting for in-person class afterwards?

Great.

